Windowed residual-minimization model reduction for dynamical systems

### 2019-09-18 main.tex
update

%\title{Model Reduction via \methodName}
\title{Windowed Least--Squares Reduced-Order Models for Dynamical Systems}

@@

\address[a]{Sandia National Laboratories,  Livermore, CA}
\begin{abstract}
This work proposes a windowed least-squares (\methodAcronym) reduced-order modeling formulation for dynamical systems. The proposed approach 
sequentially minimizes the time-continuous full-order model residual within a low-dimensional trial space over a series of time windows.
Trial spaces with spatial and space--time dimension reduction are considered. 
The stationary conditions for the resulting minimization problems are derived. Of particular interest, the stationary conditions for trial spaces with spatial dimension reduction are obtained via the Euler--Lagrange equations and comprise a coupled two-point Hamiltonian boundary value problem containing a forward and backward system. %The forward system, which is forced by an auxiliary costate, evolves the generalized coordinates of the reduced-order model. The backwards system, which is forced by the FOM residual evaluated about the ROM state, governs the dynamics of the costate.
%For spatial trial spaces, the stationary conditions for the time-continuous minimization problems are obtained 
%by deriving the associated Euler--Lagrange equations. 
Two differing methods are examined to solve the minimization problems: direct (discretize then minimize) and indirect (minimize and then discretize). Various limiting cases of the proposed method yield existing reduced-order modeling techniques, namely Galerkin, Least-Squares Petrov--Galerkin, and Space--Time Least-Squares Petrov--Galerkin. 
%By formulating the least--squares problem from the time-continuous level, the \methodAcronym\ formulation overcomes 
%the time-step sensitivity and time-scheme dependence that LSPG is subject to. 
%The stationary points of the minimization problems are obtained by solving the associated Euler-Lagrange equations. The Euler-Lagrange equations consist of a set of coupled two point boundary value problems defined on each time-slab containing a forward primal system and a backward adjoint system. %The approach is conceptually similar to the classic Least-Squares Petrov-Galerkin approach, but displays several advantages. First, by formulating the residual minimization problem at the continuous level, the sensitivity of the resulting method to the underlying time-discretization scheme is greatly reduced. 
Numerical experiments on the compressible Euler and Navier--Stokes equations demonstrate that, at an increased computational cost, the \methodAcronym\ formulation can yield more accurate and physically relevant solutions than the Galerkin and Least-Squares Petrov--Galerkin approaches.

@@

%\maketitle
\section{Introduction}
The simulation of dynamical systems plays an important role in a variety of areas in science and engineering; such examples include engineering design and control, financial analysis, and epidemiology. It is often the case, however, that (numerically) executing a single dynamical system solve is a computationally intensive process; e.g., this is the case when the state-space dimension of the dynamical system is prohibitively large and/or a wide range of time-scales need to be resolved. In such circumstances, it is common for analysts to rely on approximate models that generate approximations to the system of interest at a greatly reduced cost.

Projection-based reduced-order models (ROMs) have proved to be a promising tool for developing low-cost approximate solutions. These ROMs leverage ``snapshot" data from high-fidelity simulations and/or numerical experiments to identify a low-dimensional ``trial" space on which the state variables can be well approximated. ROM methods then seek a solution to the governing equations that lies on this low-dimensional trial space via projection or residual minimization. Efforts by a variety of research groups have led to significant advancements in ROM technologies over the past several decades, and we refer to Ref.~\cite{wilcox_benner_rev} for a review.

The construction of reduced-order models for nonlinear dynamical systems has proved challenging. The principle challenge encountered when trying to construct reduced-order models for such problems is that most existing methods lack \textit{a priori} guarantees of stability, accuracy, and convergence. %The standard Galerkin and least-squares Petrov-Galerkin methods, for instance, are both subject to exponentially growing error bounds~\cite{carlberg_lspg_v_galerkin}. 
A great deal of effort has been devoted towards developing methods that allow for more robust and accurate reduced-order models of dynamical systems. These efforts include residual-norm minimization~\cite{l1,carlberg_lspg,carlberg_gnat,legresley_1,legresley_2,legresley_3,bui_resmin_steady,bui_unsteady,rovas_thesis,carlberg_thesis,bui_thesis}, space--time methods~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC,benner_st},  ``energy-based" inner products~\cite{rowley_pod_energyproj,Kalashnikova_sand2014}, symmetry transformations~\cite{sirovich_symmetry_trans}, basis adaptation~\cite{carlberg_hadaptation,adeim_peherstorfer}, projection subspace rotations~\cite{basis_rotation}, and subgrid-scale modeling~\cite{san_iliescu_geostrophic,iliescu_pod_eddyviscosity,iliescu_vms_pod_ns,Bergmann_pod_vms,iliescu_ciazzo_residual_rom,Wang_ROM_thesis,wentland_apg,Wang:269133,San2018}. Many of these developed techniques are not mutually exclusive. Residual minimization is most relevant to the current work and is the focus of the following review.

>>>>

%\title{Model Reduction via \methodName}
%\title{Windowed least--squares reduced-order models for dynamical systems}
\title{Windowed residual-minimization model reduction for dynamical systems}

@@

\address[a]{Sandia National Laboratories,  Livermore, CA}
\begin{abstract}
This work proposes a \methodNameLower\ (\methodAcronym) model-reduction
	formulation for dynamical systems. The proposed approach minimizes the
	time-continuous full-order-model residual over a low-dimensional space--time
	trial subspace across a sequence of time windows that decompose the time
	domain of interest.  We consider two types of space--time trial subspaces
	within this framework:
	one that reduces the spatial dimension of the full-order model, and one that
	reduces both the spatial and temporal dimensions of the full-order model.
	\KTC{Why are we saying the next two sentences? These seem kind of random.
	The abstract should explain why the work is necessary; this more reads like
	a list of interesting factoids.} The stationary conditions for the resulting minimization problems are
	derived.  Of particular interest, the stationary conditions for trial spaces
	with spatial dimension reduction are obtained via the Euler--Lagrange
	equations and comprise a coupled two-point Hamiltonian boundary value
	problem containing a forward and backward system. 
%The forward system, which is forced by an auxiliary costate, evolves the generalized coordinates of the reduced-order model. The backwards system, which is forced by the FOM residual evaluated about the ROM state, governs the dynamics of the costate.
%For spatial trial spaces, the stationary conditions for the time-continuous minimization problems are obtained 
%by deriving the associated Euler--Lagrange equations. 
We consider two methods to solve the residual-minimization problems: direct
	(i.e., discretize-then-optimize) and indirect (i.e.,
	optimize-then-discretize).  Critically, we show that the proposed method
	comprises a generalization of many existing techniques and provides a
	unified framework to reason about existing model-redcution techniques. In
	particular, we show that specific instances of the proposed method lead to
	the existing reduced-order modeling techniques of Galerkin projection,
	least-squares Petrov--Galerkin (LSPG) projection, and space--time
	least-squares Petrov--Galerkin (ST-LSPG) projection. 
%By formulating the least--squares problem from the time-continuous level, the \methodAcronym\ formulation overcomes 
%the time-step sensitivity and time-scheme dependence that LSPG is subject to. 
%The stationary points of the minimization problems are obtained by solving the associated Euler-Lagrange equations. The Euler-Lagrange equations consist of a set of coupled two point boundary value problems defined on each time-slab containing a forward primal system and a backward adjoint system. %The approach is conceptually similar to the classic Least-Squares Petrov-Galerkin approach, but displays several advantages. First, by formulating the residual minimization problem at the continuous level, the sensitivity of the resulting method to the underlying time-discretization scheme is greatly reduced. 
Numerical experiments on the compressible Euler and Navier--Stokes equations demonstrate that, at an increased computational cost, the \methodAcronym\ formulation can yield more accurate and physically relevant solutions than the Galerkin and Least-Squares Petrov--Galerkin approaches.

@@

%\maketitle
\section{Introduction}

Simulating parameterized dynamical systems underpins many applications across
science and engineering, including engineering design, control, financial
analysis, and epidemiology. In many contexts, executing a dynamical-system
simulation at a single parameter instance---which entails the numerical
integration of a system of ordinary differential equations (ODEs)---incurs an
extremely large computational expense.  This occurs, for example, when the
state-space dimension is large (e.g., due to fine spatial resolution when
discretizing a partial-differential-equations problem) and/or when the number
of time instances is large (e.g., due to stability or accuracy requirements
that limit the time step).  When the application is time-critical or
many-query in nature, analysts must replace such large-scale parameterized
dynamical-system models with a low-cost approximation that makes the
application tractable.

\KTC{I reworked this. Methods like balanced truncation or IRKA do \textit{not}
use snapshot data!}

Projection-based reduced-order models (ROMs) comprise one such approximation
strategy. First, these techniques execute a computationally expensive
\textit{offline} stage that computes a low-dimensional ``trial'' subspace on
which the (parameterized) dynamical-system state can be well approximated
(e.g., by computing state ``snapshots'' at different parameter instances, by
solving Lyapunov equations). Second, these methods execute an inexpensive
\textit{online} stage during which they compute approximations to the
dynamical-system trajectory that reside on this trial subspace via a
residual-minimization or projection process.

Model reduction for linear-time-invariant systems (and other strongly
structured dynamical systems) is quite mature \cite{wilcox_benner_rev}
\KTC{Also cite classical refs on balanced truncation, Gugercin's IRKA,. Note
that I moved \cite{wilcox_benner_rev} here because that review mainly deals
with linear systems}, as system-theoretic properties (e.g., controllability,
observability, asymptotic stability, $\mathcal H_2$-optimality) can be readily
quantified and accounted for in the projection process, which yields
reduced-order models that inherit such properties.  The primary challenge for
developing reduced-order models for general nonlinear dynamical systems is
that such properties are difficult to assess for such systems. As a result, it
is difficult to develop projection methods that preserve important
dynamical-system properties in the reduced-order model, which can lead to
models that are inaccurate or unstable.  To address this, researchers have
pursued several directions that aim to imbue reduced-order models of nonlinear
dynamical systems with properties that can improve robustness and accuracy.
These
efforts include residual-minimizing
projection that imbues the ROM solution with a notion of optimality~\cite{l1,carlberg_lspg,carlberg_gnat,legresley_1,legresley_2,legresley_3,bui_resmin_steady,bui_unsteady,rovas_thesis,carlberg_thesis,bui_thesis};
space--time
projection that leads to error bounds that grow slowly in time~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC,benner_st};
``energy-based'' inner
products that ensure non-increasing entropy in the ROM solution~\cite{rowley_pod_energyproj,Kalashnikova_sand2014}; symmetry
transformations that \KTC{what?} ~\cite{sirovich_symmetry_trans},
basis-adaptation methods that improve the ROM's accuracy \textit{a posteriori}~\cite{carlberg_hadaptation,adeim_peherstorfer}, subspace
rotations that \KTC{what?} ~\cite{basis_rotation}, structure-preserving
methods that enforce conservation \cite{carlberg_conservative_rom} or
(port-)Hamiltonian/Lagrangian structure
\cite{carlberg2012spd,structurePreserveBeattie,chaturantabut2016structure,farhat2014dimensional}
\KTC{Can use more here. See my presentation} in the ROM; and subgrid-scale
modeling methods that aim to improve accuracy by addressing the closure
problem~\cite{san_iliescu_geostrophic,iliescu_pod_eddyviscosity,iliescu_vms_pod_ns,Bergmann_pod_vms,iliescu_ciazzo_residual_rom,Wang_ROM_thesis,wentland_apg,Wang:269133,San2018}.
We note that these techniques are often not mutually exclusive.
Residual-minimizing and space--time projection are the most relevant classes of
methods to the current work and comprise the focus of the following review.
\KTC{Shouldn't we review space-time also?}

\KTC{HERE}
